---
title: "DATA605 Homework 13 Univariate & Multivariate Calculus"
author: "Omar Pineda"
date: "11/29/2020"
output:
  html_document: default
  pdf_document: default
---

1. Use integration by substitution to solve the integral below.

$$\int 4e^{-7x}dx$$

Let u = -7x, du/dx = -7, and du = -7dx. By substituting we have:

$$\int -4/7* e^{u}du$$

$$=-4/7\int e^{u}du$$ 

$$F(u) = -4/7*e^{u} + C$$
$$=-4/7*e^{-7x} + C$$


2. Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of dN/dt = -3150/t^4 - 220 bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N(t) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

We have the following integral that we evaluate:

$$\int -3150/t^{-4} - 220dt$$
$$F(t) = (3150/t^{-3})/3 - 220t + C$$
When t = 1, this function is equal to 6530, and we can use this to find C for the general equation.

$$F(1) =1050/(1)^{-3} - 220(1) + C = 6530$$
$$C = 6530 - 1050 + 220 = 5700$$
So, we have the following generalized equation: 

$$1050/(t)^{-3} - 220(t) + 5700$$

3. Find the total area of the red rectangles in the figure below, where the equation of the line is f(x) = 2x - 9.

```{r}
1+3+5+7
```

4. Find the area of the region bounded by the graphs of the given equations. y = x^2 - 2x - 2, y = x + 2

First, we need to find where these two equations intersect

x^2 - 2x - 2 = x + 2

x^2 - 3x - 4 = (x - 4) (x + 1) = 0

So, these two equations intersect at x = 4 and x = -1 and we integrate the difference between these functions within this interval.

$$\int_{-1}^4 (x+2) - (x^2 - 2x - 2)dx$$

$$=\int_{-1}^4 -x^2 + 3x + 4 dx$$
$$F(x)=-x^3/3 + 3x^2/2+ 4x + C$$

Finally, we plug 4 and -1 into this equation to get the total area between the these functions.

```{r}
(-(4)^3/3 + 3*(4)^2/2 + 4*(4)) - (-(-1)^3/3 + 3*(-1)^2/2 + 4*(-1))
```

5. A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

Inventory costs will be minimized if all 110 flat irons are sold in one order since the fixed cost of $8.25 would only be incurred once.

6. Use integration by parts to solve the integral below.

$$\int ln(9x) * x^6 dx$$

Let f(x) = ln(9x), f'(x) = 1/x, g'(x) = x^6, g(x) = x^7/7

We have that 

$$\int f(x) * g'(x) dx = f(x)*g(x) - \int f'(x) * g(x) dx$$

Plugging in our functions, we get the following

$$ln(9x)*(x^7/7) - \int 1/x * x^7/7 dx $$

$$= ln(9x)*(x^7/7) - \int x^6/7 dx $$

$$= ln(9x)*(x^7/7) - x^7/49 + C $$
$$= (x^7/7)(ln(9x) - 1/7) + C $$

7. Determine whether f(x) is a probability density function on the interval [1, e^6]. If not, determine the value of the definite integral. f(x) = 1/6x

We have the following integral:

$$\int_{1}^{e^6} 1/(6x) dx$$
$$= 1/6*ln(x)$$
Evaluating this over the interval [1, e^6], we get that this is equal to 1, so we can conclude that f(x) is a probability density function on this interval.

```{r}
(1/6*log(exp(1)^6)) - (1/6*log(1))
```

