---
title: "DATA605 Homework 3 Eigenvalues and Eigenvectors"
author: "Omar Pineda"
date: "9/13/2020"
output:
  html_document: default
  pdf_document: default
---

1. Problem set 1

(1) What is the rank of the matrix A?

$$\mathbf{A} = \left[\begin{array}
{rrr}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & 3
\end{array}\right]
$$

```{r}
A <- rbind(c( 1, 2, 3, 4),
           c(-1, 0, 1, 3),
           c( 0, 1,-2, 1),
           c( 5, 4,-2, -3))
qr(A)$rank
```

(2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

The maximum rank would be n, and the mininum rank would be 1.

(3) What is the rank of matrix B?

$$\mathbf{B} = \left[\begin{array}
{rrr}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2 
\end{array}\right]
$$

```{r}
B <- rbind(c( 1, 2, 1), 
           c( 3, 6, 3), 
           c( 2, 4, 2))
qr(B)$rank
```


2. Problem set 2
Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your work. You’ll need to write out the characteristic polynomial and show your solution.

A =
1 2 3
0 4 5
0 0 6

To get the characteristic polynomial, we solve the following:

$det(A - \lambda I) = 0$

$(1 - \lambda)(4 - \lambda)(6 - \lambda) = 0$

$(4 - \lambda - 4\lambda + \lambda^2)(6 -\lambda) = 0$

$(4 - 5\lambda + \lambda^2)(6 - \lambda) = 0$

$(24 - 4\lambda - 30 \lambda + 5 \lambda^2 + 6 \lambda^2 - \lambda^3) = 0$

$- \lambda ^3 + 11 \lambda^2 - 34 \lambda + 24 = 0$

We can confirm that this is the characteristic polynomial through the following code:

```{r}
library(pracma)
a <- rbind(c(1, 2, 3), 
           c(0, 4, 5),
           c(0, 0, 6))

A <- charpoly(a, info = TRUE)
A$cp
```

Our solutions to this characteristic polynomial are the eigenvalues 6, 4, and 1.

Finally, we calculate the eigenvectors corresponding to these eigenvalues.

```{r}
ev <- eigen(a)$vectors
ev
```