---
title: "DATA605 Homework 6 Combinatorics/Conditional Prob"
author: "Omar Pineda"
date: "10/4/2020"
output:
  html_document: default
  pdf_document: default
---

1. A box contains 54 red marbles, 9 white marbles, and 75 blue marbles. If a marble is randomly selected from the box, what is the probability that it is red or blue? Express your answer as a fraction or a decimal number rounded to four decimal places.

(75+54)/138 = 129/138

2. You are going to play mini golf. A ball machine that contains 19 green golf balls, 20 red golf balls, 24 blue golf balls, and 17 yellow golf balls, randomly gives you your ball. What is the probability that you end up with a red golf ball? Express your answer as a simplified fraction or a decimal rounded to four decimal places.

20/80 = 1/4

3. A pizza delivery company classifies its customers by gender and location of residence. The research department has gathered data from a random sample of 1399 customers. The data is summarized in the table below.

What is the probability that a customer is not male or does not live with parents? Write your answer as a fraction or a decimal number rounded to four decimal places.

This is essentially all women plus any men who do not live with parents, so all participants minus the men who live with their parents.

(1399-215)/1399 = 1184/1399

4. Determine if the following events are independent. Going to the gym. Losing weight.
Answer: A) Dependent B) Independent

A) These events are dependent because P(Losing Weight | Going to the gym) != P(Losing Weight).

5. A veggie wrap at City Subs is composed of 3 different vegetables and 3 different condiments wrapped up in a tortilla. If there are 8 vegetables, 7 condiments, and 3 types of tortilla available, how many different veggie wraps can be made?

8C3 * 7C3 * 3C1 = 5880

```{r}
choose(8,3) * choose(7,3) * choose(3,1)
```

6. Determine if the following events are independent. Jeff runs out of gas on the way to work. Liz watches the evening news.
Answer: A) Dependent B) Independent

B) These are independent because P(Jeff runs out of gas on the way to work | Liz watches the evening news) = P(Jeff runs out of gas on the way to work)

7. The newly elected president needs to decide the remaining 8 spots available in the cabinet he/she is appointing. If there are 14 eligible candidates for these positions (where rank matters), how many different ways can the members of the cabinet be appointed?

Since rank matters, this is a permutation rather than a combination.

```{r}
factorial(14)/factorial(14-8)
```

8. A bag contains 9 red, 4 orange, and 9 green jellybeans. What is the probability of reaching into the bag and randomly withdrawing 4 jellybeans such that the number of red ones is 0, the number of orange ones is 1, and the number of green ones is 3? Write your answer as a fraction or a decimal number rounded to four decimal places.

a. The number of ways to choose 0 jellybeans out of 9 total red jellybeans AND choose 1 jellybean out of 4 total orange jellybeans AND choose 3 jellybeans out of 9 total green jellybeans is equal to 9C0 * 4C1 * 9C3 = 336.

```{r}
choose(9,0) * choose(4,1) * choose(9,3)
```

b. There are 22 total jellybeans, so there are 22C4 = 7,315 different possible ways to choose 4 jellybeans.

```{r}
choose(22,4)
```

Finally, we divide a) by b) to get the probability that this draw would happen.

```{r}
336/7315
```

9. Evaluate the following expression.
11!
7!

```{r}
factorial(11)/factorial(7)
```

10. Describe the complement of the given event: 67% of subscribers to a fitness magazine are over the age of 34.

33% of subcribers to a fitness magazine are NOT over the age of 34

11. If you throw exactly three heads in four tosses of a coin you win $97. If not, you pay me $30. 

Step 1. Find the expected value of the proposition. Round your answer to two decimal places.

The probability of a win, meaning tossing a coin four times AND having 3 of those tosses be heads, is equal to 0.25.

```{r}
prob_win <- (1/2)^4 * choose(4,3)
prob_win
```

The probability of losing is then 0.75.

```{r}
prob_lose <- 1 - prob_win
```

The value of the proposition is thus 1.75, which is pretty low but still in the positive.

```{r}
value <- 97*prob_win - 30*prob_lose
value
```

Step 2. If you played this game 559 times how much would you expect to win or lose? (Losses must be entered as negative.)

If we played this game 559 times, we would expect to win a total of $978.25.

```{r}
559*value
```

12. Flip a coin 9 times. If you get 4 tails or less, I will pay you $23. Otherwise you pay me $26.

Step 1. Find the expected value of the proposition. Round your answer to two decimal places.

The probability of a win, meaning tossing a coin 9 times AND having 4 OR 3 OR 2 OR 1 OR 0 tosses be tails, is equal to 0.5.

```{r}
prob_win2 <- (1/2)^9 * (choose(9,4) + choose(9,3) + choose(9,2) + choose(9,1) + choose(9,0))
prob_win2
```

The probability of losing is then 0.5.

```{r}
prob_lose2 <- 1 - prob_win2
```

The value of the proposition is thus -1.5, so we would end up losing.

```{r}
value2 <- 23*prob_win2 - 26*prob_lose2
value2
```

Step 2. If you played this game 994 times how much would you expect to win or lose? (Losses must be entered as negative.)

If we played this game 994 times, we would expect to lose a total of $1,491.

```{r}
994*value2
```

13. The sensitivity and specificity of the polygraph has been a subject of study and debate for years. A 2001 study of the use of polygraph for screening purposes suggested that the probability of detecting a liar was .59 (sensitivity) and that the probability of detecting a "truth teller" was .90 (specificity). We estimate that about 20% of individuals selected for the screening polygraph will lie.

Here we construct a confusion matrix for actual values vs predicted values.

```{r}
prob_actual_lie <- 0.2
prob_actual_truth <- 0.8
prob_lie_detected <- 0.59
prob_truth_detected <- 0.9

prob_lie_detected_lie <- prob_actual_lie*prob_lie_detected
prob_lie_detected_truth <- prob_actual_lie - prob_lie_detected_lie

prob_truth_detected_truth <- prob_actual_truth*prob_truth_detected
prob_truth_detected_lie <- prob_actual_truth-prob_truth_detected_truth

conf_mat <- data.frame(predict_lie = c(prob_lie_detected_lie, prob_truth_detected_lie),
                       predict_truth = c(prob_lie_detected_truth, prob_truth_detected_truth)
                       )
row.names(conf_mat) <- c('actual_lie', 'actual_truth')

conf_mat
```

a. What is the probability that an individual is actually a liar given that the polygraph detected him/her as such? (Show me the table or the formulaic solution or both.)

Using the above table, we get:

```{r}
0.118/(0.118+0.08)
```

b. What is the probability that an individual is actually a truth-teller given that the polygraph detected him/her as such? (Show me the table or the formulaic solution or both.)

Using the table, we get:

```{r}
0.72/(0.72+0.082)
```

c. What is the probability that a randomly selected individual is either a liar or was identified as a liar by the polygraph? Be sure to write the probability statement.

This probability is equal to P(liar) + P(predict_lie) = (0.118+0.082) + (0.118+0.080) - 0.118 (We subtract 0.118 because there is overlap between the two probabilities). 

```{r}
(0.118+0.082) + (0.118+0.080) - 0.118
```

Weekly discussion post:

3.15 A computing center has 3 processors that receive n jobs, with the jobs assigned to the processors purely at random so that all of the 3^n possible assignments are equally likely. Find the probability that exactly one processor has no jobs.

For this question, we need to find the probability that one processor has 0 jobs AND the other two processors have a non-zero number of jobs. First, we choose one of the three processors at random to have 0 jobs, and this is given by 3C1. We then have 2^n possible combinations of how the jobs can be assigned between the remaining two processors. However, since we only want to have one processor with no jobs, we need to remove two additional options from 2^n: when either of the two remaining processors have all n jobs and the other has 0 jobs. This gives us 2^n - 2 possible assignments. So, we have 3C1 * (2^n -2) possible assignments where one processor has 0 jobs and the other two have a non-zero number of jobs.

Finally, we divide by the total number of possible assignments (3^n) to get (3C1 * (2^n -2))/3^n as the probability that exactly one processor has no jobs.