---
title: "DATA605 Homework 9 Central Limit Theorem & Generating Functions"
author: "Omar Pineda"
date: "10/25/2020"
output:
  html_document: default
  pdf_document: default
---

11. The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by Yn on the nth day of the year. Finn observes that the differences Xn = Yn+1 − Yn appear to be independent random variables with a common distribution having mean µ = 0 and variance σ^2 = 1/4. If Y1 = 100, estimate the probability that Y365 is

(a) ≥ 100.

```{r}
pnorm(100, mean = 100, sd = sqrt(1/4*364), lower.tail = F)
```

(b) ≥ 110.

```{r}
pnorm(110, mean = 100, sd = sqrt(1/4*364), lower.tail = F)
```

(c) ≥ 120.

```{r}
pnorm(120, mean = 100, sd = sqrt(1/4*364), lower.tail = F)
```

2. Calculate the expected value and variance of the binomial distribution using the moment generating function.

We have the following:

$$g(t)=\sum_{i=1}^∞e^{tx_i}px_i = (pe^t + q) ^n$$
$$g'(0)=n(pe^0 + q)^{n-1}pe^0 = np$$
$$g''(0)=n(n-1)p^2 + np$$

So, the mean is:

$$\mu = np$$
and the variance is:

$$\sigma^2 = g''(0) - g'(0)^2 = [n(n-1)p^2 + np] - (np)^2 = np(1-p)$$

3. Calculate the expected value and variance of the exponential distribution using the moment generating function.

We have the following:

$$g(t) = \int_{-∞}^∞ e^{tx}f(x)dx = \int_{0}^∞ e^{tx}\lambda e^{-\lambda x}dx = \lambda/(\lambda - t)$$

$$g^n(0) = n!/\lambda^n$$
$$g'(0) = 1/\lambda$$
$$g''(0) = 2/\lambda^2$$
So, the mean is:

$$\mu = 1/\lambda$$
and the variance is:

$$\sigma^2 = g''(0) - g'(0)^2 = 2/\lambda^2 - (1/\lambda)^2 = 1/\lambda^2$$

Weekly discussion post:

Chapter 9.2, Exercise 1

A die is rolled 24 times. Use the Central Limit Theorem to estimate the probability that

(a) the sum is greater than 84.

We have that E(X) = (1*1/6) + (2*1/6) + (3*1/6) + (4*1/6) + (5*1/6) + (6*1/6) = 7/2 and V(X) = 1/6 * ((1−7/2)^2 + (2−7/2)^2 + (3−7/2)^2 + (4−7/2)^2 + (5−7/2)^2 + (6−7/2)^2) = 35/12. 

For 24 rolls, E(X) = 24*7/2 = 84, and V(X) = 24*35/12 = 70. So, σ = sqrt(70) = 8.37. The probability that the sum of 24 die rolls is greater than 84 is P(sum of 24 rolls > 84) = P(sum of 24 rolls > (84.5 - 84)/8.37) = 

```{r}
1 - pnorm((84.5-84)/8.37)
```

(b) the sum is equal to 84.

P(sum of 24 rolls = 84) = P(sum of 24 rolls < (84.5 - 84)/8.37) - P(sum of 24 rolls < (83.5 - 84)/8.37) =

```{r}
pnorm((84.5-84)/8.37) - pnorm((83.5-84)/8.37)
```