---
title: "DATA605 Homework 2 Trace, Determinant, Factorization"
author: "Omar Pineda"
date: "9/6/2020"
output:
  pdf_document: default
  html_document: default
---

1. Problem set 1

(1) Show that A^T x A != A x A^T in general. (Proof and demonstration.)

Let

$$\mathbf{A} = \left[\begin{array}
{rrr}
A11 & A12 \\
A21 & A22 
\end{array}\right]
$$
then we have that

$$\mathbf{A^T} = \left[\begin{array}
{rrr}
A11 & A21 \\
A12 & A22 
\end{array}\right]
$$

For A^t x A, we have a 2 x 2 matrix with a value of (A11 x A11 + A12 X A12) in the [1,1] position. In A x A^T we have a 2 x 2 matrix with a value of (A11 x A11 + A21 x A21).

(A11 x A11 + A12 X A12) != (A11 x A11 + A21 x A21) unless A12 = A21, therefore A^T x A != A x A^T in general.

(2) For a special type of square matrix A, we get A^T x A = A x A^T. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix). 

A^T x A = A x A^T would hold true if A was a symmetric matrix as is the identity matrix.

2. Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. 

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

In this function, we assume that matrix A can indeed be decomposed.

```{r}
LU_decomposition <- function(A){
  U <- A
  #calculate U by replacing all positions below its diagonal with zeros
  for (j in 1:(ncol(U)-1)) {
    for (i in (j+1):nrow(U)) {
      if (U[i,j]!=0) {
        U[i,] <- U[i, ] - U[j, ]*(U[i,j]/U[j,j])
      }
    }
  }
  #calculate L by solving for L_1 in U_1 x L_1 = A_1
  A_1 <- solve(A)
  U_1 <- solve(U)
  L_1 <- solve(U_1, A_1)
  L <- solve(L_1)
  #return L and U and have their values include 2 decimals
  list ("L" = round(L, 2), "U" = round(U,2))
}
```

We then test our LU decomposition function with example 1 from the following website: http://mathonline.wikidot.com/the-lu-decomposition-of-a-matrix-examples-1. Matrix A was successfully decomposed.

$$\mathbf{A} = \left[\begin{array}
{rrr}
3 & 1 \\
4 & 2 
\end{array}\right]
$$

```{r}
A <- rbind(c(3, 1),
                c(4, 2))
LU <- LU_decomposition(A)
print(LU)
```