---
title: "DATA605 Homework 7 Imp. Distributions / EX / VARX"
author: "Omar Pineda"
date: "10/11/2020"
output:
  html_document: default
  pdf_document: default
---

1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.

The cumulative distribution function of the minimum of a function uniform on an interval from a to b is denoted by f(x) = 1 - ((b-x)/(b-a))^n.

Here, we want the distribution function of Y which is the minimum of Xi's that are each uniform on the internal 1 to k. So, Y is uniform on the interval 1 to k as well and we plug this interval into our above equation to get f(y) = 1 - ((k-y)/(k-1))^n as our distribution.

2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

This is the probability:

```{r}
p <- 1/10
1-sum(dgeom(0:8, p))
```

This is the mean:

```{r}
1/p
```

This is the standard deviation:

```{r}
sqrt(((1-p)/p^2))
```

b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

This is the probability:

```{r}
l <- 1/10
pexp(8, l, lower.tail = F)
```

This is the mean and the standard deviation:

```{r}
1/l
```

c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

This is the probability:

```{r}
pbinom(0, 8, p)
```

This is the mean:

```{r}
8 * p
```

This is the standard deviation:

```{r}
sqrt(8*p*(1-p))
```

d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson. 

This is the probability:

```{r}
ppois(0, 8*p)
```

This is the mean:

```{r}
8*p
```

This is the standard deviation:

```{r}
sqrt(8*p)
```

Weekly discussion post:

Chapter 5.1 Exercise 8

If a coin is tossed a sequence of times, what is the probability that the first head will occur after the fifth toss, given that it has not occurred in the first two tosses?

We want the probability that the first 5 tosses are tails given that the first 2 tosses are tails. This is denoted by P((T,T,T,T,T)|(T,T)) = (1-p)^5 / (1-p)^2 = (1-p)^3. Plugging in 0.5 for p in our equation, we get that this probability is equal to 0.125.

```{r}
(1-.5)^3
```