---
title: "DATA605 Homework 11 Regression Analysis in R"
author: "Omar Pineda"
date: "11/15/2020"
output:
  html_document: default
  pdf_document: default
---

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

First, we load any necessary libraries.

```{r}
library(dplyr)
library(ggplot2)
```

We then preview our cars dataset.

```{r}
head(cars)
```

Here we visualize a scatterplot of stopping distance by speed.

```{r}
cars %>% ggplot(aes(x=speed, y=dist)) + geom_point() + geom_smooth(method="lm") + labs(x="Speed", y="Stopping Distance", title="Stopping Distance by Speed")
```

We then generate our linear model. It appears that speed is a statistically significant predictor for stopping distance, and it also explains about 65% of the variability in stopping speed.

```{r}
m1 <- lm(dist ~ speed, data=cars)
summary(m1)
```

Lastly, we conduct some residual analysis to determine whether this linear model is appropriate.

### Inference: (1) Linearity and (2) Constant Variability

(1) There does not appear to be a pattern in the residuals of the linear regression, so the relationship between Speed and Stopping Distance is linear.

(2) Based on our scatter plot of the residuals, we appear to have constant variability.

```{r residuals, eval=TRUE}
plot(m1$residuals ~ cars$speed)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
```

### Inference: (3) Nearly normal residuals

(3) The distribution of our model's residuals is slightly skewed but still nearly normal. The linear model is a pretty good start but there may be better options out there for modeling stopping distance.

```{r hist-res, eval=TRUE}
hist(m1$residuals)
```