---
title: "DATA605 Homework 15 Functions of Several Variables"
author: "Omar Pineda"
date: "12/12/2020"
output:
  html_document: default
  pdf_document: default
---

1. Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary. ( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

y = 4.2571x - 14.8

```{r}
x_vals <- c(5.6, 6.3, 7, 7.7, 8.4)
y_vals <- c(8.8, 12.4, 14.8, 18.2, 20.8)
m1 <- lm(y_vals ~ x_vals)
summary(m1)
```

2. Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.

f(x,y) = 24x - 6xy^2 - 8y^3

First we find the partial derivates:

$$f_x(x,y) = 24 - 6y^2$$
$$f_y(x,y) = -12xy - 24y^2$$
We then solve for y by setting fx = 0

$$f_x(x,y) = 24 - 6y^2 = 0, y= 2, -2$$
and plug in -2 and 2 for y to find their respective x values when fy = 0

$$f_y(x,-2) = -12xy - 24y^2 = -12x(-2) - 24(-2)^2 = 24x - 96 = 0, x = 4$$
$$f_y(x,2) = -12xy - 24y^2 = -12x(2) - 24(2)^2 = -24x - 96 = 0, x = -4$$
Using the original formula to find z, we have

$$f(4,-2) = 24(4) - 6(4)(-2)^2 - 8(-2)^3 = 96 - 96 + 64 = 64$$
and

$$f(-4,2) = 24(-4) - 6(-4)(2)^2 - 8(2)^3 = -96 + 96 - 64 = -64$$

So our critical points are (4,-2,64), (-4,2,-64)

3. A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the "house" brand and 40 + 11x - 23y units of the "name" brand.

Step 1. Find the revenue function R ( x, y ).

$$R(x, y) = x(81 - 21x + 17y) + y(40 + 11x - 23y) = 81x - 21x^2 + 17xy + 40y + 11xy - 23y^2 = - 21x^2 - 23y^2 + 28xy + 81x + 40y$$

Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

```{r}
x <- 2.3
y <- 4.1
- 21*x^2 - 23*y^2 + 28*x*y + 81*x + 40*y
```

4. A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by C(x, y) = 1/6 x^2 + 1/6 y^2 + 7x + 25y + 700, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

First we find the partial derivates:

$$C_x(x,y) = 1/3*x + 7$$

$$C_y(x,y) = 1/3*y + 25$$
After setting Cy and Cx equal to 0, we can set them equal to each other in order to get y in terms of x.

$$1/3*y + 25 = 1/3*x + 7 , y = 3(1/3*x - 18) = x - 54 = y$$
We want the total number of units to equal 96, so we can use this equation to plug in y in terms of x and solve for x

$$x + y = 96 = x + (x - 54) = 2x - 54, x = 75$$
$$ y = x - 54 = 75 - 54 = 21$$
So, 75 units should be produced in LA and 21 units should be produced in Denver in order to minimize the total weekly cost.

5. Evaluate the double integral on the given region.

$$\int \int_R (e^{8x+3y})dA$$
Where R: 2 <= x <= 4 and 2 <= y <= 4

Write your answer in exact form without decimals.

$$\int_{2}^4 \int_{2}^4 (e^{8x+3y})dydx$$
First we evaluate the inner integral

$$ \int_{2}^4 (1/3*e^{8x+3*4}) - (1/3*e^{8x+3*2})dx = 1/3 * \int_{2}^4 e^{8x+12} - e^{8x+6}dx $$
Then we solve for the outer integral

$$ 1/3 * [(1/8*e^{8*4+12} - 1/8*e^{8*4+6}) -  (1/8*e^{8*2+12} - 1/8*e^{8*2+6})] = 1/24 * [e^{44} - e^{38} - e^{28} + e^{22}]$$
